---
title: Confidence Intervals of the Mean
subtitle: How to be confident about your uncertainty
format: 
    clean-revealjs
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
author:
  - name: Connor Smith
    orcid: 0000-0002-3955-4348
    email: connor.smith@mq.edu.au
    affiliations: Macquarie University
date: last-modified
---

```{r}
set.seed(123)
```

## An Initial Guess (Point Estimation)

### Lecture Attendance

At the beginning of semester, I guessed that the average lecture attendance would be $100$ students. *(Data collected from Weeks 1 to 7)*

i. This was my best guess, but I was uncertain about it.
ii. I have a few tricks up my sleeve to help me be more confident about my estimate.



## A Better Guess (Interval Estimation)

### Lecture Attendance

I collected data from previous semesters and used this information to improve my estimate.

i. The smallest average attendance was $81.1$ students.

ii. The largest average attendance was $118.9$ students.

So, to improve my estimate, I can say that the average lecture attendance will probably be between $81.1$ and $118.9$ students.


:::{.callout-important}
## Actual Data

I have been collecting data on lecture attendance for the first seven weeks of semester. The average attendance was $95.2$ students.
:::

## Outline

### Confidence Intervals

If we are estimating a population parameter $\mu$, we can use a confidence interval to express our uncertainty about the estimate.

We can write this confidence interval as

$$
\operatorname{CI} = \Big(L(X),U(X)\Big)
$$

where $L(X)$ is the lower bound and $U(X)$ is the upper bound.

:::{.callout-note}
## Using our data

Since $L(X)$ and $U(X)$ are a function of the random sample $X$, we will use our data to calculate this interval.

:::

## Desirable Properties

### Confidence Intervals

<br>

1. The confidence interval contains the true population parameter with a specified probability.

<br>

2. The interval is the shortest possible for a specified probability.

## General Form

### Confidence Intervals

<br>

We can write the general form of a confidence interval as

$$
\text{sample statistic} \pm \text{margin of error}
$$

where the margin of error is a function of the sample size $n$ and the sample standard deviation $s$.

## Example

### Confidence Intervals

In tabletop games, dice are used to generate random numbers. After purchasing many different dice, I have a bag of $n=100$ dice.

:::: {.columns}

::: {.column width="70%"}
```{r}
#| code-fold: true
#| echo: true

N <- 100
sides <- c(4, 6, 8, 10, 12, 20)
dice <- sample(sides, N, replace = TRUE)
```

If I take a sample of size $n=5$, I can calculate the sample mean $\bar{x}$.

```{r}
#| code-fold: true
#| echo: true

n1 <- 5
sample1_dice <- sample(dice, n1, replace = FALSE)
mean1_dice <- mean(sample1_dice)
```
:::

::: {.column width="30%"}
```{r}
#| code-fold: true

knitr::kable(data.frame(sample = 1:n1, sides = sample1_dice), caption = "First Sample", format = "html", col.names = c("Sample", "Sides"))
```

:::

::::



## Example Cont.

### Confidence Intervals



## Actual Count

### Confidence Intervals

:::: {.columns}

::: {.column width="40%"}

```{r}
#| code-fold: true

knitr::kable(table(dice), caption = "Dice in the bag", format = "html", col.names = c("Dice", "Count"))
```
:::

::: {.column width="60%"}
Let's see how we did. The table shows the number of each type of dice in the dice bag.

We can see that the population mean is $\mu_{\text{sides}} =$ `r mean(dice)`.

:::

::::


## How confident are we?

### Confidence Intervals

The confidence level of the interval is the probability that the interval contains the true population parameter.

$$
P(L(X) \leqslant \mu \leqslant U(X)) = 1 - \alpha
$$

where $\alpha$ (usually $0.05$) is the significance level. We can also think about this as

$$
P(\mu < L(X)) + P(\mu > U(X)) = \alpha
$$



## Improving our estimate

### Confidence Intervals

The confidence interval is a function of the sample size $n$ and the sample standard deviation $s$ (if the population standard deviation in unknown).

:::{.incremental}

i. As the sample size increases, the confidence interval becomes narrower.
ii. As the sample standard deviation decreases, the confidence interval becomes narrower.

:::

## A Bag of Dice

### Confidence Intervals

```{r}
#| code-fold: true

N <- 100
sides <- c(4, 6, 8, 10, 12, 20)
dice <- sample(sides, N, replace = TRUE)
```

Suppose we have a bag of `r N` dice. 

## Sample Mean

### Confidence Intervals

My population in this example is the bag of dice. I want to estimate the average number of sides on the dice, $\mu_{\text{sides}}$.

```{r}
#| code-fold: true
#| echo: true

n1 <- 5
sample1_dice <- sample(dice, n1, replace = FALSE)
mean1_dice <- mean(sample1_dice)
```

If I take a sample fo size $n =$ `r n1`, I can calculate the sample mean $\bar{x} =$ `r mean1_dice`.

## Sample Mean Again

### Confidence Intervals

```{r}
sample2_dice <- sample(dice, n1, replace = FALSE)
mean2_dice <- mean(sample2_dice)

knitr::kable(data.frame(sample1_dice, sample2_dice), caption = "Sample of dice sides", format = "html", col.names = c("Sample 1", "Sample 2"))
```

Lets repeat this process again. We take another sample of size $n =$ `r n1` and calculate the sample mean $\bar{x} =$ `r mean2_dice`.

## Other Population Parameters

### Confidence Intervals

We can also use confidence intervals to estimate other population parameters. We might also be interested in:

i. The population variance $\sigma^2$,
ii. The population standard deviation $\sigma$, and
iii. The population proportion $p$.

This list is in no way exhaustive.

# Assessment

## Individual Work

### Formative Assessment

Describe a situation where using a confidence interval instead of a point estimate would be more effective in communicating a topic of interest. (Choose a topic that is related to your major)

- Links to cohort specific topics,

- Use case scenarios, and

- Identify context specific scenarios to shape future assessments.

## Group Work

### Formative Assessment

The group work will be conducted in class the week after the lecture:

(a) Provide a subjective $95\%$ confidence interval for a tutorial specific fact. Discuss.

(b) Repeat step (a) for a new fact but include a point estimate. Discuss.

(c) Create a $95\%$ confidence interval for a new fact using actual data and CLT. Discuss.

## Assessment

### Summative Assessment



## Final Exam

### Summative Assessment




# Pedagogical Approach

## Choice of Approach

### Pedagogical Approach

**Behaviourism:** Due to the content heavy nature of statistics, the initial focus will be on rote learning and memorisation of key concepts.

**Constructivism:** Explored through Tutorials and Workshops, students will be able to construct their own understanding of the material.

**Reflective:** Through individual and group work, students will be able to reflect on their learning and identify areas for improvement.

**Socratic:** The end goal but finding a practical implementation is difficult.


# Educational Excellence

## Initiative 01

### Educational Excellence

Implementation of "Minute Papers" to improve student engagement and feedback.

1. Identify gaps in knowledge and effectiveness of teaching for a specific cohort.

2. Share common thoughts and feelings with the student cohort.

3. Use data to adapt tutorial activities to improve learning outcomes (and address student concerns).

:::{.callout-tip}
## Sydney in 2032 Strategy

*To drive a transformational experience, we will improve how we learn from data and insights from our students, staff and partners to ensure teaching contributes more effectively to desired learning outcomes*

:::

## Initiative 02

### Educational Excellence

Implementation of Universal Design for Learning (UDL) principles to improve inclusivity and accessibility.

1. Develop a more inclusive learning environment for all students.

2. Provide a range of teaching and assessment methods to cater for diverse learning styles.

3. Ensure that all students have the opportunity to succeed, regardless of their background or circumstances.

:::{.callout-tip}
## Sydney in 2032 Strategy

*By ensuring our academic and teaching support staff are highly skilled in inclusive teaching practices, we will be better equipped to support students on their learning journeys.*

:::